


## R Basics
Main reference here is @Gro2014 and @Wic2017.

### Data structures

#### Vectors
Vector is one of the basic data structure in R. It is created by `c()` function. You may store any data types in it. The data type can be checked by `typeof` function.

You may compare it with the `list` in Python.

```{r}
vec1 <- c(1, 2, 3, 5)
typeof(vec1)
```

Note that vector index starts from 1 in R, while list index starts from 0 in Python.

```{r}
vec1[1]
```
When slicing with vectors, don't forget to use `c()`.

```{r}
vec1[c(2, 3)]
```

You may use `length()` function to get its length.

```{r}
length(vec1)
```

#### matrices


```{r}
m <- matrix(c(1,2,3,4,5,6), nrow=2)
m[1, ]
```

#### `data.frame`


```{r}
df <- data.frame(face = c("ace", "two", "six"),
                 suit = c("clubs", "clubs", "clubs"),
                 value = c(1, 2, 3))
df
```


### Load data

#### build-in datasets
R has many build-in datasets. You may use `data()` to see all of them. Here are a few common datasets.

- `mtcars`: Motor Trend Car Road Tests: The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models)

```{r}
data(mtcars)
```

- `iris`: iris data set gives the measurements in centimeters of the variables sepal length, sepal width, petal length and petal width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

```{r}
data(iris)
```

- `ToothGrowth`: ToothGrowth data set contains the result from an experiment studying the effect of vitamin C on tooth growth in 60 Guinea pigs. 

```{r}
data(ToothGrowth)
```

- `PlantGrowth`: Results obtained from an experiment to compare yields (as measured by dried weight of plants) obtained under a control and two different treatment condition.

```{r}
data(PlantGrowth)
```

- `USArrests`: This data set contains statistics about violent crime rates by us state.

```{r}
data(USArrests)
```

#### Read from files
The build-in `read.csv()` function can directly read `.csv` file into a data.frame.


::: {#exm-}
We use the file `yob1880.txt` from Chapter 5 here. Put the file in the working folder and run the following code.

```{r}
#| eval: false
df <- read.csv('yob1880.txt', header = FALSE)
head(df)
```

```{r}
#| echo: false
df <- read.csv('assests/datasets/yob1880.txt', header = FALSE) 
```
We may also manually assign columns names.

```{r}
names(df) <- c('name', 'sex', 'counts')
head(df)
```
:::

### Flow control

#### `for` loop


::: {#exm-}

```{r}
for (x in 1:10){
    print(x)
}
```

:::


#### `if-else` 


::: {#exm-}

```{r}
a <- 200
b <- 33

if (b > a) {
  print("b is greater than a")
} else if (a == b) {
  print("a and b are equal")
} else {
  print("a is greater than b")
}
```

:::


#### Functions


::: {#exm-}

```{r}
myfunction <- function() {
    die <- 1:6
    sum(die)
}

myfunction()
```

:::